---
title: "Lab 12"
author: "Drew Hoffman, Tanner Ball, Burke Parent, David Washburn, Brandon Nelsen"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gapminder Questions

Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

<Question 1>

<Question 2>

* Briefly describe the gapminder dataset.

<answer>

* Explain why your questions are important. One of the questions should be answered with the two-sample mean permutation test, and the other with the correlation permutation test.

<Explanation 1>

<Explanation 2>

* Finish the function skeletons to implement the two-sample permuation test and the correlation permutation test. Run a two-sample mean permutation test. Run a correlation permutation test.

Two-Sample Permutation Function:

```{r, eval = F}

perm_mean <- function(perms = 1000, all_values, n_A)
{
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "nA" and "nB" respectively
    rand_vector <- sample(all_values)
    group_a <- rand_vector[1:n_A]
    group_b <- rand_vector[(n_A+1):length(all_values)]
    
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    mean_a <- mean(group_a)
    mean_b <- mean(group_b)
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    diff_means <- mean_a - mean_b
    perm_mean_diffs[i] <- diff_means
    
  }
  
  # Step 5:
  # Return the permuted mean differences vector
  return(perm_mean_diffs)
  
}
```

Correlation Permutation Function:

```{r, eval = F}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  
}
```

* Create visualizations of the statistical tests.

<Visualization 1>

<Visualization 2>

* Make data-based conclusions about the results. Do you believe there are differences among the groups you are comparing? 

<Conclusion>

* Generate percentiles, summary statistics (max, min), and probabilites (p-values) to justify your conclusions.

<Justification>

* Summarize each individual's contribution to the lab.

Tanner:

Drew: 

Brandon:

David:

Burke:

