---
title: "Lab 12"
author: "Drew Hoffman, Tanner Ball, Burke Parent, David Washburn, Brandon Nelsen"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(gapminder)
UgandaUSA <- filter(gapminder_unfiltered, country == "Uganda" | country == "United States", year >= 1977)
lifeExpcombined <- c(UgandaUSA$lifeExp)
```

# Gapminder Questions

Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

*Question one:*

Has life expectancy in Uganda changed (percentage-wise) more than the United States since 1977?

*Question Two:*

IS GDP per capita correlated with population?

*Briefly describe the gapminder dataset.*

The gapminder dataset was created to give access of easy to digest data about various socioeconomic statistics in each country, it is made to educate others and help bring to light some of those socioeconomic issues around the globe.

*Explain why your questions are important. One of the questions should be answered with the two-sample mean permutation test, and the other with the correlation permutation test.*

Our first question is important because we want to see if the development of a random country we picked in Africa is happening at an accelerated pace due possibly to the rise of new technology or the advancement of aid to third world countries. This can help us understand the pace at which African and third world countries in general can become first-world, if possible.

Our second question is important because it can help us look at future trends in population growth based on GDP, if a country with high GDP has more population growth, it can help a country better prepare for the influx in population or if a country is growing rapidly it can help a country prepare for a growth in GDP.

*Finish the function skeletons to implement the two-sample permuation test and the correlation permutation test. Run a two-sample mean permutation test. Run a correlation permutation test.*

Two-Sample Permutation Function:

```{r, eval = F}

perm_mean <- function(perms = 1000, all_values, n_A)
{
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "values" into disjoint 
    # groups of size "nA" and "nB" respectively
    rand_vector <- sample(all_values)
    group_a <- rand_vector[1:n_A]
    group_b <- rand_vector[(n_A+1):length(all_values)]
    
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    mean_a <- mean(group_a)
    mean_b <- mean(group_b)
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1
    diff_means <- mean_a - mean_b
    perm_mean_diffs[i] <- diff_means
    
  }
  
  # Step 5:
  # Return the permuted mean differences vector
  return(perm_mean_diffs)
  
}
```

Correlation Permutation Function:

```{r, eval = F}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  cor_coeffs <- numeric(perms)

  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    mix_up <- sample(y)

    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    coefficient <- cor(x, mix_up)
    cor_coeffs[i] <- coefficient
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(cor_coeffs)
}


```

* Create visualizations of the statistical tests.

```{r}
mean_dif <- tibble(test = perm_mean(1000, lifeExpcombined, length(lifeExpcombined) / 2))
UGN <- filter(UgandaUSA, country == "Uganda")
USA <- filter(UgandaUSA, country == "United States")
UGNdiff <- 51.542 - 50.35
USAdiff <- 78.242 - 73.38
USdiff_UGNdiff <- USAdiff - UGNdiff

ggplot(mean_dif) + geom_histogram(aes(test)) + geom_vline(xintercept = USdiff_UGNdiff)
```

<Visualization 2>

* Make data-based conclusions about the results. Do you believe there are differences among the groups you are comparing? 



* Generate percentiles, summary statistics (max, min), and probabilites (p-values) to justify your conclusions.

<Justification>

* Summarize each individual's contribution to the lab.

Tanner:

Drew: 

Brandon:

David:

Burke:

